Stability concerns in random matrices

### 2018-04-19 ms.Rmd
Add Figure 2

@@

**Figure 1**

```{r, eval = TRUE, echo = FALSE, fig.height = 8}
make_gammas <- function(nn = 10, distribution = 1, mn = 10, sdd = 1){
    if(distribution == 0){
        dat          <- rep(x = mn, times = nn);
    }
    if(distribution == 1){
        dat          <- rnorm(n = nn, mean = mn, sd = sdd);
        dat[dat < 0] <- -dat[dat < 0];
    }
    if(distribution == 2){
        dat          <- runif(n = nn, min = 0, max = 20);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 3){
        dat          <- rexp(n = nn);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 4){
        dat          <- 1 - (sdd * rexp(n = nn));
        dat          <- dat - min(dat);
        dat          <- dat - mean(dat) + mn;
        dat[dat < 0] <- -dat[dat < 0];
    }
    if(distribution == 5){
        dat          <- rbeta(n = nn, shape1 = 0.5, shape2 = 0.5);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    if(distribution == 6){
        dat          <- rgamma(n = nn, shape = 2, scale = 2);
        dat          <- sdd * ((dat - mean(dat)) / sd(dat));
        dat          <- dat + mn;
    }
    return(dat);
}
show_gammas <- function(nn = 1000000){
    y1 <- make_gammas(nn = nn, distribution = 1);
    y2 <- make_gammas(nn = nn, distribution = 2);
    y3 <- make_gammas(nn = nn, distribution = 3);
    y4 <- make_gammas(nn = nn, distribution = 4);
    y5 <- make_gammas(nn = nn, distribution = 5);
    y6 <- make_gammas(nn = nn, distribution = 6);
    par(mfrow = c(3, 2), oma = c(6, 6, 1, 1), mar = c(4, 0.5, 0.5, 0.5));
    hist(y1, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y2, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y3, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y4, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y5, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    hist(y6, breaks = 1000, yaxt = "n", main = "", xlab = "", cex.axis = 1.5);
    box();
    mtext(expression(paste("Component ",gamma[i]," value")),
          outer=TRUE,side=1,line=3.0,cex=2);
    mtext(expression(paste("Relative frequency")),
          outer=TRUE,side=2,line=2.5,cex=2);
}
show_gammas();
```

@@

>>>>

```{r, echo = FALSE}
source(file = "R/sim_mat.R");
source(file = "R/plot_figs.R");
```

@@

**Figure 1: Distributions of component response rate ($\boldsymbol{\gamma}$) values in complex systems.** The stabilities of simulated complex systems with these $\gamma$ distributions are compared to otherwise identical complex systems with a fixed component response rate of $\gamma = 1$ across different system sizes ($S$; i.e., component numbers) given a unit $\gamma$ standard deviation ($\sigma_{\gamma} = 1$). Distributions are as follows: (a) uniform, (b) exponential, (c) beta ($\alpha = 0.5$ and $\beta = 0.5$), and (d) gamma ($k = 2$ and $\theta = 2$). Each panel shows 1 million randomly generated $\gamma$ values.

```{r, eval = TRUE, echo = FALSE, fig.height = 8}
show_gammas();
```

@@

```{r, eval = TRUE, echo = FALSE}
eig_cloud <- eigen_cloud(sp = 20, iters = 1000);
Pr_D0     <- sum(eig_cloud$d0[,1] < 0) / length(eig_cloud$d0[,1]);
Pr_D1     <- sum(eig_cloud$d1[,1] < 0) / length(eig_cloud$d1[,1]);
Pr_D2     <- sum(eig_cloud$d2[,1] < 0) / length(eig_cloud$d2[,1]);
Pr_D3     <- sum(eig_cloud$d3[,1] < 0) / length(eig_cloud$d3[,1]);
Pr_D4     <- sum(eig_cloud$d4[,1] < 0) / length(eig_cloud$d4[,1]);
```

**Figure 2: Distributions of eigenvalues across different distributions of component response rate  ($\boldsymbol{\gamma}$) values.** Panels show 100000 random communities for $S = 20$ and $C = 1$, where off-diagonal elements are randomly sampled from a standard normal distribution $N(0, 1)$, diagonal elements are $-1$, and each row of elements is multiplied by a random scalar value $\gamma$ sampled from one of four distributions (black points in all sub-panels):  **a**, uniform $\Pr(\gamma) = 1/\sqrt{12}$ for $x \in [0, \sqrt{12}]$. **b**, exponential $\Pr(\gamma) = e^{-\lambda}$ for $\lambda = 1$. **c**, beta $\Pr(\gamma) = \frac{x^{\alpha = 1}(1-x)^{\beta - 1}}{B(\alpha, \beta)}$, where $B(\alpha, \beta) = \frac{\Gamma(\alpha)\Gamma(\beta)}{\Gamma(\alpha + \beta)}$ for $\alpha = 0.5$ and $\beta = 0.5$. **d**, gamma $\Pr(\gamma) = \frac{1}{\Gamma(k)\theta^{k}}x^{k-1}e^{\frac{-x}{\theta}}$ for $k = 2$ and $\theta = 2$. Eigenvalues associated with each $\gamma$ distribution are compared with otherwise identical communities in which $\gamma = 1$ (pink points in all sub-panels). In all distributions, $SD(\gamma) = 1$; note that $\gamma$ distributions cannot be mean standardised without changing the sign of some community matrix elements, but the sign of eigenvalues (and therefore community stability) is unaffected by multiplying elements by any scalar value. Proportions of real components of eigenvalues (x-axis) below 0 are (a) `r Pr_D1`, (b) `r Pr_D2`, (c) `r Pr_D3`, (d) `r Pr_D4`, as compared to `r Pr_D0` when $\gamma = 1$.

```{r, eval = TRUE, echo = FALSE, fig.height = 8}
eig_plot(ecl = eig_cloud);
```
